---
title: "Bulk Uploading Mammal Data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Bulk Uploading Mammal Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The purpose of the `bulkUploader` package is to help data managers to bring their data into the [Neotoma Paleoecology Database](https://neotomadb.org).  Neotoma is a highly structured database that contains information on a large number of data types, including diatoms, ostracodes, pollen and mammal fossils.

Neotoma is really a "database of databases", and contains X constituent databases.  To help data managers include their data in Neotoma, we have wrapped Neotoma package functions around tools to help manage, validate and upload their data into Neotoma.

```{r setup, results='hide', message=FALSE}
library(bulkUploader)
devtools::install_github('NeotomaDB/neotoma2')
library(neotoma2)
library(purrr)
library(dplyr)
```

## Managing your Data

The easiest data to enter into Neotoma is the publication data and contact data.  From a given database, or data resource, we want to check that information is accurate.  For contacts, unless we have ORCIDs for all users, we're stuck hand-checking information.  For publication records we can use the DOI as a "gold standard" for publication information, since CrossRef manages this data for us, and provides a unique identifier that we can check against Neotoma's information.

### Publication Data

However your publication data is organized, it must be aligned with Neotoma's so that it can be included in the database.  The Neotoma `publications` table uses the following fields in the `publication` object:

| Field | Description |
=========== | =================
| publicationtype | The type of publication |
| publicationid | A unique Neotoma publication identifier |
| articletitle | The title of the document |
| year | The publishing year where available |
| journal | The journal (if available) |
| volume | Journal volume |
| issue | Journal issue |
| pages | Pages over which the item spans |
| citation | A plain text citation, generally APA format. |
| doi | The article DOI |
| author |  A list of `contacts` |

Users can directly add new information into Neotoma using the `neotoma2::add_publication()` function, or they can use `scrape_dois()`, a function in the `bulkUploader` package, to validate publication information and to add additional metadata, where such metadata does not already exist.

```{r scrapedois, results='hide'}

# Note that `miomapcsv` is a data element contained within the `bulkUploader` package 

data <- data(miomapcsv)

# Ideally this code would be massaged.  The issue here is that there is title and journal information in several places.  It's not the nicest way of doing things.

pubList <- formatPubs(miomapcsv, 
                      author='Author', 
                      year='Year', 
                      title=c('Title', 'Book.Title'))

outputs <- scrape_dois(pubList[1:10,], 
            n = 3, 
            savefile = 'miotest.RDS', 
            restore = FALSE)

results <- map(outputs, get_cr_results)
```

From this output we have a `list` object called `outputs`, with a length equal to the number of publications in the original data file.  We also have a `list` called `results`.  The `list` has a length equal to the length of `outputs` and is made up of individual `data.frame`s, with columns `score` and `delta`, showing the quality of each match returned by the CrossRef API, and the quality of the next matches.  This is useful to understand the quality of the matches.  High quality matches with have a high `diff` score between the first and second matches, and a generally high `score` value.

The `score` gives the approximate match strength to a publication in CrossRef.  The `delta` lets us know what the score of the next closest match is.  Because we need to work with the CrossRef API in a reasonable way, this matching tends to be slow.

Once this script runs we then have a set of publications for the constituent database, and for each described publication we have 0 or more potential matches.  To generate a simple match we want to create some form of citation.  We can do this using the very simple helper function `makeCite()`.

```{r reviewPubs}
miocites <- makeCite(miomapcsv$Author,
                     miomapcsv$Year,
                     miomapcsv$Title,
                     miomapcsv$Journal)

crossCites <- outputs %>% map(cr_to_neotoma)
```

Now, for each publication we have a set of possible matches, and scores.  We need to prompt the user to select the best match:

```{r getMatches, results='hide'}
bestMatch <- matchPubs(cross = crossCites,
                       pubs = miocites,
                       results = results)

```
