---
title: "Bulk Uploading Mammal Data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Bulk Uploading Mammal Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The purpose of the `bulkUploader` package is to help data managers to bring their data into the [Neotoma Paleoecology Database](https://neotomadb.org).  Neotoma is a highly structured database that contains information on a large number of data types, including diatoms, ostracodes, pollen and mammal fossils.

Neotoma is really a "database of databases", and contains X constituent databases.  To help data managers include their data in Neotoma, we have wrapped Neotoma package functions around tools to help manage, validate and upload their data into Neotoma.

## Managing your Data

The easiest data to enter into Neotoma is the publication data and contact data.  From a given database, or data resource, check 

### Publication Data

However your publication data is organized, it must be aligned with Neotoma's so that it can be included in the database.  The Neotoma `publications` table uses the following fields:

```{r setup}
library(bulkUploader)
library(neotoma2)
```

```{r}

results <- data.frame(score=rep(NA, nrow(miomapcsv)), delta = NA)
outputs <- list()

for (i in 1:nrow(miomapcsv)) {
  container = ifelse(is.na(miomapcsv$Journal[i]),
                     miomapcsv$Book.Title[i],
                     miomapcsv$Journal[i])
  if(is.na(container)) {
    message("Row ", i, " is missing a title.")
  } else {
    outputs[[i]] <- scrape_dois(title = miomapcsv$Title[i],
                        year = miomapcsv$Year[i])
    if (length(test) > 0) {
      results$score[i] <- test[[1]]$score
      if (length(test) > 1) {
        results$delta[i] <- test[[1]]$score - test[[2]]$score
      }
      message("Tested article ", i)
    } else {
      message("Couldn't get a match for ", i)
    }
  }
}
```

